---
title: "Draft"
output: html_notebook
---

# loading and creating dataset for modeling
The following scripts loads the data set from SE and CN sites and segment the data set into vehicle-vehicle interaction and vehicle-vru interactions. For the CN dataset, we manually assign the interaction ID so that only interaction where all SMoS measurements are available are kept. For each type of the interaction, datasets for running both BM and POT is created.
```{r message=FALSE, warning=FALSE, include=FALSE}
source('preprocessing.R')

# load injury severity models
PIS0 <- function(speed){return(1/(1+exp(5.261 - 0.104*speed))) } # Eq.6 
PIS1 <- function(speed,age){return(1/(1+exp(5.15 - 0.101*speed - 0.042*age))) } # Eq.10 

# MAIS 3+ for VRU Lubbe et al (2022)
PIS2 <- function(speed,age){return(1/(1+exp(6.19 - 0.078*speed - 0.038*age))) } # Table 2, for ped
PIS3 <- function(speed,age){return(1/(1+exp(7.47 - 0.079*speed - 0.047*age))) } # Table 2, for cyclist

injury.df <- data.frame(speed = seq(0,80,0.5)) %>%
  mutate(InjuryP = sapply(speed, PIS0))
```

# assigning the indicators to global variables.
# TTC and Speed_TTC 
```{r,echo=FALSE}
Dat.SE <- SE_TTC
Dat.CN <- CN_TTC_A
x0.1 <- 0
x0.2 <- 0   # crash boundary 
source("scripts/global_vars.R")

Qc(x0.1,Dat.CN$prox,u1)
Qc(x0.2,Dat.SE$prox,u2)
Qc(x0.1,Dat.CN$prox,v.1)
Qc(x0.2,Dat.SE$prox,v.2)
```
# iTTC and Speed_TTC
```{r,echo=FALSE}
Dat.SE <- SE_iTTC
Dat.CN <- CN_iTTC_A
x0.1 <- 1
x0.2 <- 1
source("scripts/global_vars.R")
Qc(x0.1,Dat.CN$prox,u1)
Qc(x0.2,Dat.SE$prox,u2)
Qc(x0.1,Dat.CN$prox,v.1)
Qc(x0.2,Dat.SE$prox,v.2)
```


## BPOT gumbel
```{r,echo=FALSE}

# source("scripts/BPOT_logistics.R")

logistic.result <- runExecutionBPOT(Dat.CN = Dat.CN,Dat.SE = Dat.SE,model="log",thres1 = c(u1,v1),
    thres2 = c(u2,v2))
# plots

scale_factor <- max(logistic.result$plot.df.1$ConditionalD)/max(injury.df$InjuryP)

ggplot(logistic.result$plot.df.1,aes(x=speed,y=ConditionalD)) + 
  geom_line(aes(colour = "CN")) + 
  geom_line(data = logistic.result$plot.df.2,aes(x=speed,y=ConditionalD,colour = "SE")) +
  geom_line(data = injury.df,aes(x=speed,y=InjuryP * scale_factor) ) +
  geom_vline(xintercept = v1, linetype = "dashed", color = "red") +
  geom_vline(xintercept = v2, linetype = "dashed", color = "blue") +
  annotate("text", x = 12, y = 0.015, 
            label = paste("u=", round(v2,3)), color = "blue") +
  annotate("text", x = 20, y = 0.02, 
            label = paste("u=", round(v1,3)), color = "red")+
  scale_y_continuous(
    name = "denisty",  # primary y-axis label
    sec.axis = sec_axis(~ . / scale_factor, name = "Injury prob")  # secondary y-axis
  ) +
  scale_colour_manual(name = "Site", values = c("CN" = "red", "SE" = "blue")) +
  labs(x = "Speed (km/h)", y = "f(y|TTC<0)") +
  theme(panel.grid.major = element_line(colour = "gray91"),
    panel.grid.minor = element_line(colour = "gray88"),
    panel.background = element_rect(fill = "white",
        colour = "white", linetype = "solid"),
    plot.background = element_rect(linetype = "solid"))
ggsave("plots/impact_speed_density_TTC_BPOT_gumbel.png")

cat("logistic dependence: thresholds CN:",logistic.result$M.1$threshold,"\n")
cat("logistic dependence: parameters SE:",logistic.result$M.2$threshold,"\n")
cat("logistic dependence: parameters CN:",logistic.result$M.1$estimate,"\n")
confint(logistic.result$M.1)
cat("logistic dependence: parameters SE:",logistic.result$M.2$estimate,"\n")
confint(logistic.result$M.2)
cat("logistic dependence: Deviance CN:",deviance(logistic.result$M.1),"\n")
cat("logistic dependence: Deviance SE:",deviance(logistic.result$M.2),"\n")
cat("logistic dependence: Crash prob CN:",logistic.result$Pcrash.1,"\n")
cat("logistic dependence: Crash prob SE:",logistic.result$Pcrash.2,"\n")
cat("logistic dependence: Severe Crash prob CN:",
    pbTvevd(q1=x0.1,q2=40,dep=logistic.result$M.1$estimate[5],thres=c(u1,v1),model="log",
        eta=logistic.result$M.1$nat[1]/logistic.result$M.1$n,
        mar1=c(logistic.result$M.1$estimate[1],logistic.result$M.1$estimate[2]),
        mar2=c(logistic.result$M.1$estimate[3],logistic.result$M.1$estimate[4]),tail.type=2),"\n")
cat("logistic dependence: Severe Crash prob SE:",
    pbTvevd(q1=x0.2,q2=40,dep=logistic.result$M.2$estimate[5],thres=c(u2,v2),model="log",
        eta=logistic.result$M.2$nat[1]/logistic.result$M.2$n,
        mar1=c(logistic.result$M.2$estimate[1],logistic.result$M.2$estimate[2]),
        mar2=c(logistic.result$M.2$estimate[3],logistic.result$M.2$estimate[4]),tail.type=2),"\n")
cat("logistic dependence: INjury prob CN:",
    Injury.from_c_bivariate(logistic.result$plot.df.1$speed,EVmodel=logistic.result$M.1,
                            severity = PIS0,x0=x0.1,PX=logistic.result$Pcrash.1),"\n")
cat("logistic dependence: INjury prob SE:",
    Injury.from_c_bivariate(logistic.result$plot.df.2$speed,EVmodel=logistic.result$M.2,
                            severity = PIS0,x0=x0.2,PX=logistic.result$Pcrash.2),"\n")

```


## BPOT HR
```{r,echo=FALSE}
hr.result <- runExecutionBPOT(Dat.CN = Dat.CN,Dat.SE = Dat.SE,model="hr",thres1 = c(u1,v1),
    thres2 = c(u2,v2))
# source("scripts/BPOT_HR.R")
# plots

scale_factor <- max(hr.result$plot.df.1$ConditionalD)/max(injury.df$InjuryP)

ggplot(hr.result$plot.df.1,aes(x=speed,y=ConditionalD)) + 
  geom_line(aes(colour = "CN")) + 
  geom_line(data = hr.result$plot.df.2,aes(x=speed,y=ConditionalD,colour = "SE")) +
  geom_line(data = injury.df,aes(x=speed,y=InjuryP * scale_factor) ) +
  geom_vline(xintercept = v1, linetype = "dashed", color = "red") +
  geom_vline(xintercept = v2, linetype = "dashed", color = "blue") +
  annotate("text", x = 12, y = 0.015, 
            label = paste("u=", round(v2,3)), color = "blue") +
  annotate("text", x = 20, y = 0.02, 
            label = paste("u=", round(v1,3)), color = "red")+
  scale_y_continuous(
    name = "denisty",  # primary y-axis label
    sec.axis = sec_axis(~ . / scale_factor, name = "Injury prob")  # secondary y-axis
  ) +
  scale_colour_manual(name = "Site", values = c("CN" = "red", "SE" = "blue")) +
  labs(x = "Speed (km/h)", y = "f(y|TTC<0)") +
  theme(panel.grid.major = element_line(colour = "gray91"),
    panel.grid.minor = element_line(colour = "gray88"),
    panel.background = element_rect(fill = "white",
        colour = "white", linetype = "solid"),
    plot.background = element_rect(linetype = "solid"))
ggsave("plots/impact_speed_density_TTC_BPOT_HR.png")

cat("Husler and Reiss dependence: thresholds CN:",hr.result$M.1$threshold,"\n")
cat("Husler and Reiss dependence: parameters SE:",hr.result$M.2$threshold,"\n")
cat("Husler and Reiss dependence: parameters CN:",hr.result$M.1$estimate,"\n")
confint(hr.result$M.1)
cat("Husler and Reiss dependence: parameters SE:",hr.result$M.2$estimate,"\n")
confint(hr.result$M.2)
cat("Husler and Reiss dependence: Deviance CN:",deviance(hr.result$M.1),"\n")
cat("Husler and Reiss dependence: Deviance SE:",deviance(hr.result$M.2),"\n")
cat("Husler and Reiss dependence: Crash prob CN:",hr.result$Pcrash.1,"\n")
cat("Husler and Reiss dependence: Crash prob SE:",hr.result$Pcrash.2,"\n")
cat("Husler and Reiss dependence: Severe Crash prob CN:",
    pbTvevd(q1=x0.1,q2=40,dep=hr.result$M.1$estimate[5],thres=c(u1,v1),model="hr",
        eta=hr.result$M.1$nat[1]/hr.result$M.1$n,
        mar1=c(hr.result$M.1$estimate[1],hr.result$M.1$estimate[2]),
        mar2=c(hr.result$M.1$estimate[3],hr.result$M.1$estimate[4]),tail.type=2),"\n")
cat("Husler and Reiss dependence: Severe Crash prob SE:",
    pbTvevd(q1=x0.2,q2=40,dep=hr.result$M.2$estimate[5],thres=c(u2,v2),model="hr",
        eta=hr.result$M.2$nat[1]/hr.result$M.2$n,
        mar1=c(hr.result$M.2$estimate[1],hr.result$M.2$estimate[2]),
        mar2=c(hr.result$M.2$estimate[3],hr.result$M.2$estimate[4]),tail.type=2),"\n")
cat("Husler and Reiss dependence: INjury prob CN:",
    Injury.from_c_bivariate(hr.result$plot.df.1$speed,EVmodel=hr.result$M.1,
                            severity = PIS0,x0=x0.1,PX=hr.result$Pcrash.1),"\n")
cat("Husler and Reiss dependence: INjury prob SE:",
    Injury.from_c_bivariate(hr.result$plot.df.2$speed,EVmodel=hr.result$M.2,PX = hr.result$Pcrash.2,
                            severity = PIS0,x0=x0.2),"\n")
```


## BPOT ct
```{r,echo=FALSE}

ct.result <- runExecutionBPOT(Dat.CN = Dat.CN,Dat.SE = Dat.SE,model="ct",thres1 = c(u1,v1),
    thres2 = c(u2,v2))
# source("scripts/BPOT_CT.R")
# plots

scale_factor <- max(ct.result$plot.df.1$ConditionalD)/max(injury.df$InjuryP)

ggplot(ct.result$plot.df.1,aes(x=speed,y=ConditionalD)) + 
  geom_line(aes(colour = "CN")) + 
  geom_line(data = ct.result$plot.df.2,aes(x=speed,y=ConditionalD,colour = "SE")) +
  geom_line(data = injury.df,aes(x=speed,y=InjuryP * scale_factor) ) +
  geom_vline(xintercept = v1, linetype = "dashed", color = "red") +
  geom_vline(xintercept = v2, linetype = "dashed", color = "blue") +
  annotate("text", x = 12, y = 0.015, 
            label = paste("u=", round(v2,3)), color = "blue") +
  annotate("text", x = 20, y = 0.02, 
            label = paste("u=", round(v1,3)), color = "red")+
  scale_y_continuous(
    name = "denisty",  # primary y-axis label
    sec.axis = sec_axis(~ . / scale_factor, name = "Injury prob")  # secondary y-axis
  ) +
  scale_colour_manual(name = "Site", values = c("CN" = "red", "SE" = "blue")) +
  labs(x = "Speed (km/h)", y = "f(y|TTC<0)") +
  theme(panel.grid.major = element_line(colour = "gray91"),
    panel.grid.minor = element_line(colour = "gray88"),
    panel.background = element_rect(fill = "white",
        colour = "white", linetype = "solid"),
    plot.background = element_rect(linetype = "solid"))
ggsave("plots/impact_speed_density_TTC_BPOT_CT.png")

cat("Coles and Tawn dependence: thresholds CN:",ct.result$M.1$threshold,"\n")
cat("Coles and Tawn dependence: parameters SE:",ct.result$M.2$threshold,"\n")
cat("Coles and Tawn dependence: parameters CN:",ct.result$M.1$estimate,"\n")
confint(ct.result$M.1)
cat("Coles and Tawn dependence: parameters SE:",ct.result$M.2$estimate,"\n")
confint(ct.result$M.2)
cat("Coles and Tawn dependence: Deviance CN:",deviance(ct.result$M.1),"\n")
cat("Coles and Tawn dependence: Deviance SE:",deviance(ct.result$M.2),"\n")
cat("Coles and Tawn dependence: Crash prob CN:",ct.result$Pcrash.1,"\n")
cat("Coles and Tawn dependence: Crash prob SE:",ct.result$Pcrash.2,"\n")
cat("Coles and Tawn dependence: Severe Crash prob CN:",
    pbTvevd(q1=x0.1,q2=40,alpha=ct.result$M.1$estimate[5],beta=ct.result$M.1$estimate[6],thres=c(u1,v1),model="ct",
        eta=ct.result$M.1$nat[1]/ct.result$M.1$n,
        mar1=c(ct.result$M.1$estimate[1],ct.result$M.1$estimate[2]),
        mar2=c(ct.result$M.1$estimate[3],ct.result$M.1$estimate[4]),tail.type=2),"\n")
cat("Coles and Tawn dependence: Severe Crash prob SE:",
    pbTvevd(q1=x0.2,q2=40,alpha=ct.result$M.2$estimate[5],beta=ct.result$M.2$estimate[6],thres=c(u2,v2),model="ct",
        eta=ct.result$M.2$nat[1]/ct.result$M.2$n,
        mar1=c(ct.result$M.2$estimate[1],ct.result$M.2$estimate[2]),
        mar2=c(ct.result$M.2$estimate[3],ct.result$M.2$estimate[4]),tail.type=2),"\n")
cat("Coles and Tawn dependence: INjury prob CN:",
    Injury.from_c_bivariate(ct.result$plot.df.1$speed,EVmodel=ct.result$M.1,
                            severity = PIS0,x0=x0.1,PX=ct.result$Pcrash.1),"\n")
cat("Coles and Tawn dependence: INjury prob SE:",
    Injury.from_c_bivariate(ct.result$plot.df.2$speed,EVmodel=ct.result$M.2,
                            PX = ct.result$Pcrash.2,severity = PIS0,x0=x0.2),"\n")
```


## BPOT nonparametric 
```{r,echo=FALSE}
source("scripts/BPOT_nonpar.R")

scale_factor <- max(plot.df.1$ConditionalD)/max(injury.df$InjuryP)

ggplot(plot.df.1,aes(x=speed,y=ConditionalD)) + 
  geom_line(aes(colour = "CN")) + 
  geom_line(data = plot.df.2,aes(x=speed,y=ConditionalD,colour = "SE")) +
  geom_line(data = injury.df,aes(x=speed,y=InjuryP * scale_factor) ) +
  geom_vline(xintercept = v1, linetype = "dashed", color = "red") +
  geom_vline(xintercept = v2, linetype = "dashed", color = "blue") +
  annotate("text", x = 12, y = 0.015, 
            label = paste("u=", round(v2,3)), color = "blue") +
  annotate("text", x = 20, y = 0.02, 
            label = paste("u=", round(v1,3)), color = "red")+
  scale_y_continuous(
    name = "denisty",  # primary y-axis label
    sec.axis = sec_axis(~ . / scale_factor, name = "Injury prob")  # secondary y-axis
  ) +
  scale_colour_manual(name = "Site", values = c("CN" = "red", "SE" = "blue")) +
  labs(x = "Speed (km/h)", y = "f(y|TTC<0)") +
  theme(panel.grid.major = element_line(colour = "gray91"),
    panel.grid.minor = element_line(colour = "gray88"),
    panel.background = element_rect(fill = "white",
        colour = "white", linetype = "solid"),
    plot.background = element_rect(linetype = "solid"))
ggsave("plots/impact_speed_density_TTC_BPOT_HR.png")

cat("Nonparametric approach: thresholds CN:",M.1$threshold,"\n")
cat("Nonparametric approach: parameters SE:",M.2$threshold,"\n")
cat("Nonparametric approach: parameters CN:",M.1$estimate,"\n")
confint(M.1)
cat("Nonparametric approach: parameters SE:",M.2$estimate,"\n")
confint(M.2)
cat("Nonparametric approach: Deviance CN:",deviance(M.1),"\n")
cat("Nonparametric approach: Deviance SE:",deviance(M.2),"\n")
cat("Nonparametric approach: Crash prob CN:",Pcrash.1,"\n")
cat("Nonparametric approach: Crash prob SE:",Pcrash.2,"\n")
cat("Nonparametric approach: Severe Crash prob CN:",
    pbTvevd(q1=x0.1,q2=40,dep=M.1$estimate[5],thres=c(u1,v1),model="hr",
        eta=M.1$nat[1]/M.1$n,
        mar1=c(M.1$estimate[1],M.1$estimate[2]),
        mar2=c(M.1$estimate[3],M.1$estimate[4]),tail.type=2),"\n")
cat("Husler and Reiss dependence: Severe Crash prob SE:",
    pbTvevd(q1=x0.2,q2=40,dep=M.2$estimate[5],thres=c(u2,v2),model="hr",
        eta=M.2$nat[1]/M.2$n,
        mar1=c(M.2$estimate[1],M.2$estimate[2]),
        mar2=c(M.2$estimate[3],M.2$estimate[4]),tail.type=2),"\n")
cat("Husler and Reiss dependence: INjury prob CN:",
    Injury.from_c_bivariate(plot.df.1,EVmodel=M.1,severity = PIS0,x0=x0.1),"\n")
cat("Husler and Reiss dependence: INjury prob CN:",
    Injury.from_c_bivariate1(plot.df.1,EVmodel=M.1,severity = PIS0,x0=x0.1,PX=Pcrash.1),"\n")
cat("Husler and Reiss dependence: INjury prob SE:",
    Injury.from_c_bivariate(plot.df.2,EVmodel=M.2,severity = PIS0,x0=x0.2),"\n")
cat("Husler and Reiss dependence: INjury prob SE:",
    Injury.from_c_bivariate1(plot.df.2,EVmodel=M.2,PX = Pcrash.2,severity = PIS0,x0=x0.2),"\n")
```


## CBPOT gumbel
```{r,echo=FALSE}
gumbel.result <- runExecutionCBPOT(Cop.Dat.1, Cop.Dat.2, 
                                   copula=list(gumbelCopula(),rotCopula(gumbelCopula(),flip=c(FALSE,TRUE))),
                                   P2.1 = Conseq.1,P2.2 = Conseq.2,
                                   Qcrash.1 = Qcrash.1, Qcrash.2 = Qcrash.2)
# source("scripts/CBPOT_gumbel.R")


gof.1 <- gofCopula(gumbel.result$Cop.1@copula, Cop.dat.1, method = "Sn",estim.method='mpl',verbose=FALSE,
          N = 1000)
gof.2 <- gofCopula(gumbel.result$Cop.2@copula, Cop.dat.2, method = "Sn",estim.method='mpl',verbose=FALSE,
          N = 1000)

#
scale_factor <- max(gumbel.result$plot.dfQ.2$ConditionalD)/max(injury.df$InjuryP)

ggplot(gumbel.result$plot.dfQ.2,aes(x=speed,y=ConditionalD)) + 
  geom_line(aes(colour = "SE")) + 
  geom_line(data = gumbel.result$plot.dfQ.1,aes(x=speed,y=ConditionalD,colour = "CN")) +
  geom_line(data = injury.df,aes(x=speed,y=InjuryP * scale_factor) ) +
  scale_y_continuous(
    name = "denisty",  # primary y-axis label
    sec.axis = sec_axis(~ . / scale_factor, name = "Injury prob")  # secondary y-axis
  ) +
  scale_colour_manual(name = "Site", values = c("CN" = "red", "SE" = "blue")) +
  labs(title = "Probability density of impact speed",
       x = "Speed (km/h)", y = "f(y|TTC<0)") +
  theme(panel.grid.major = element_line(colour = "gray91"),
    panel.grid.minor = element_line(colour = "gray88"),
    panel.background = element_rect(fill = "white",
        colour = "white", linetype = "solid"),
    plot.background = element_rect(linetype = "solid"))
ggsave("plots/impact_speed_density_TTC_CBPOT_gumbel.png")



cat("logistic dependence: threshold CN:",v.1,"\n")
cat("logistic dependence: threshold SE:",v.2,"\n")
cat("logistic dependence: parameters prox CN:","\n")
ci(POT.1,type="parameter")
cat("logistic dependence: parameters prox SE:","\n")
ci(POT.2,type="parameter")
cat("logistic dependence: parameters conseq CN:",Conseq.1$estimate,"\n")
confint(Conseq.1)
cat("logistic dependence: parameters conseq SE:",Conseq.2$estimate,"\n")
confint(Conseq.2)
cat("logistic dependence: copula param CN:","lb",gumbel.result$Cop.1@estimate-1.96 * sqrt(gumbel.result$Cop.1@var.est),
    "estim",gumbel.result$Cop.1@estimate,"ub",gumbel.result$Cop.1@estimate+1.96 * sqrt(gumbel.result$Cop.1@var.est),"\n")
cat("logistic dependence: copula param SE:","lb",gumbel.result$Cop.2@estimate-1.96 * sqrt(gumbel.result$Cop.2@var.est),
    "estim",gumbel.result$Cop.2@estimate,"ub",gumbel.result$Cop.2@estimate+1.96 * sqrt(gumbel.result$Cop.2@var.est),"\n")

cat("logistic dependence: Deviance CN:",-2*gumbel.result$Cop.1@loglik,"\n")
cat("logistic dependence: Deviance SE:",-2*gumbel.result$Cop.2@loglik,"\n")
cat("logistic dependence: GoF CN:","test statistics",gof.1$statistic,"p-value",gof.1$p.value,"\n")
cat("logistic dependence: GoF SE:","test statistics",gof.2$statistic,"p-value",gof.2$p.value,"\n")
cat("logistic dependence: Crash prob CN in Q:",Qcrash.1,"\n")
cat("logistic dependence: Crash prob SE in Q:",Qcrash.2,"\n")
# cat("logistic dependence: Severe Crash prob CN in Q:",pMvdc(c(x0.1,40),CM0.1),"\n")
# cat("logistic dependence: Severe Crash prob SE in Q:",pMvdc(c(x0.2,40),CM0.2),"\n")
cat("logistic dependence: Injury prob CN:",
    Injury.from_cQ_bivariate(gumbel.result$Cop.1@copula,PIS0,x0.1.un,Qcrash.1*0.3,Conseq.1),"\n") # 0.3 is the threshold probability
cat("logistic dependence: Injury prob SE:",
    Injury.from_cQ_bivariate(gumbel.result$Cop.2@copula,PIS0,x0.2.un,Qcrash.2*0.2,Conseq.2),"\n")
```

## CBPOT Clayton

```{r,echo=FALSE}
clayton.result <- runExecutionCBPOT(Cop.Dat.1, Cop.Dat.2, 
                                   copula=list(claytonCopula(),rotCopula(claytonCopula(),flip=c(TRUE,FALSE))),
                                   P2.1 = Conseq.1,P2.2 = Conseq.2,
                                   Qcrash.1 = Qcrash.1, Qcrash.2 = Qcrash.2)
#source("scripts/CBPOT_clayton.R")

gof.1 <- gofCopula(clayton.result$Cop.1@copula, Cop.dat.1, method = "Sn",estim.method='mpl',verbose=FALSE,
          N = 1000)
gof.2 <- gofCopula(clayton.result$Cop.2@copula, Cop.dat.2, method = "Sn",estim.method='mpl',verbose=FALSE,
          N = 1000)

#
scale_factor <- max(clayton.result$plot.dfQ.2$ConditionalD)/max(injury.df$InjuryP)

ggplot(clayton.result$plot.dfQ.2,aes(x=speed,y=ConditionalD)) + 
  geom_line(aes(colour = "SE")) + 
  geom_line(data = clayton.result$plot.dfQ.1,aes(x=speed,y=ConditionalD,colour = "CN")) +
  geom_line(data = injury.df,aes(x=speed,y=InjuryP * scale_factor) ) +
  scale_y_continuous(
    name = "denisty",  # primary y-axis label
    sec.axis = sec_axis(~ . / scale_factor, name = "Injury prob")  # secondary y-axis
  ) +
  scale_colour_manual(name = "Site", values = c("CN" = "red", "SE" = "blue")) +
  labs(title = "Probability density of impact speed",
       x = "Speed (km/h)", y = "f(y|TTC<0)") +
  theme(panel.grid.major = element_line(colour = "gray91"),
    panel.grid.minor = element_line(colour = "gray88"),
    panel.background = element_rect(fill = "white",
        colour = "white", linetype = "solid"),
    plot.background = element_rect(linetype = "solid"))
ggsave("plots/impact_speed_density_TTC_CBPOT_clayton.png")



cat("clayton dependence: threshold CN:",v.1,"\n")
cat("clayton dependence: threshold SE:",v.2,"\n")
cat("clayton dependence: parameters prox CN:","\n")
ci(POT.1,type="parameter")
cat("clayton dependence: parameters prox SE:","\n")
ci(POT.2,type="parameter")
cat("clayton dependence: parameters conseq CN:",Conseq.1$estimate,"\n")
confint(Conseq.1)
cat("clayton dependence: parameters conseq SE:",Conseq.2$estimate,"\n")
confint(Conseq.2)
cat("clayton dependence: copula param CN:","lb",clayton.result$Cop.1@estimate-1.96 * sqrt(clayton.result$Cop.1@var.est),
    "estim",clayton.result$Cop.1@estimate,"ub",clayton.result$Cop.1@estimate+1.96 * sqrt(clayton.result$Cop.1@var.est),"\n")
cat("clayton dependence: copula param SE:","lb",clayton.result$Cop.2@estimate-1.96 * sqrt(clayton.result$Cop.2@var.est),
    "estim",clayton.result$Cop.2@estimate,"ub",clayton.result$Cop.2@estimate+1.96 * sqrt(clayton.result$Cop.2@var.est),"\n")

cat("clayton dependence: Deviance CN:",-2*clayton.result$Cop.1@loglik,"\n")
cat("clayton dependence: Deviance SE:",-2*clayton.result$Cop.2@loglik,"\n")
cat("clayton dependence: GoF CN:","test statistics",gof.1$statistic,"p-value",gof.1$p.value,"\n")
cat("clayton dependence: GoF SE:","test statistics",gof.2$statistic,"p-value",gof.2$p.value,"\n")
cat("clayton dependence: Crash prob CN in Q:",Qcrash.1,"\n")
cat("clayton dependence: Crash prob SE in Q:",Qcrash.2,"\n")
#cat("clayton dependence: Severe Crash prob CN in Q:",pMvdc(c(x0.1,40),CM0.1),"\n")
#cat("clayton dependence: Severe Crash prob SE in Q:",pMvdc(c(x0.2,40),CM0.2),"\n")
cat("clayton dependence: Injury prob CN:",
    Injury.from_cQ_bivariate(clayton.result$Cop.1@copula,PIS0,x0.1.un,Qcrash.1*0.3,Conseq.1),"\n")
cat("clayton dependence: Injury prob SE:",
    Injury.from_cQ_bivariate(clayton.result$Cop.2@copula,PIS0,x0.2.un,Qcrash.2*0.2,Conseq.2),"\n")
```

## CBPOT TawnT1

```{r,echo=FALSE}
source("scripts/CBPOT_T1.R")


gof.1 <- gofEVCopula_VC2(t1.result$Cop.1@copula, Cop.dat.1, method = "ml",estimator = 'CFG',ties.method = "average",
          N = 1000)
gof.2 <- gofEVCopula_VC2(t1.result$Cop.2@copula, Cop.dat.2, method = "ml",estimator = 'CFG',ties.method = "average",
          N = 1000)


#
scale_factor <- max(t1.result$plot.dfQ.2$ConditionalD)/max(injury.df$InjuryP)

ggplot(t1.result$plot.dfQ.2,aes(x=speed,y=ConditionalD)) + 
  geom_line(aes(colour = "SE")) + 
  geom_line(data = t1.result$plot.dfQ.1,aes(x=speed,y=ConditionalD,colour = "CN")) +
  geom_line(data = injury.df,aes(x=speed,y=InjuryP * scale_factor) ) +
  scale_y_continuous(
    name = "denisty",  # primary y-axis label
    sec.axis = sec_axis(~ . / scale_factor, name = "Injury prob")  # secondary y-axis
  ) +
  scale_colour_manual(name = "Site", values = c("CN" = "red", "SE" = "blue")) +
  labs(title = "Probability density of impact speed",
       x = "Speed (km/h)", y = "f(y|TTC<0)") +
  theme_minimal()
ggsave("plots/impact_speed_density_TTC_CBPOT_T1.png")



cat("T1 dependence: threshold CN:",v.1,"\n")
cat("T1 dependence: threshold SE:",v.2,"\n")
cat("T1 dependence: parameters prox CN:","\n")
ci(POT.1,type="parameter")
cat("T1 dependence: parameters prox SE:","\n")
ci(POT.2,type="parameter")
cat("T1 dependence: parameters conseq CN:",Conseq.1$estimate,"\n")
confint(Conseq.1)
cat("T1 dependence: parameters conseq SE:",Conseq.2$estimate,"\n")
confint(Conseq.2)
cat("T1 dependence: copula param CN:","lb",t1.result$Cop.1@estimate[1]-1.96 * sqrt(t1.result$Cop.11$se),
    "estim",Cop.1@estimate[1],"ub",Cop.1@estimate[1]+1.96 * sqrt(Cop.11$se),"\n",
    Cop.1@estimate[2]-1.96 * sqrt(Cop.11$se2),
    "estim",Cop.1@estimate[2],"ub",Cop.1@estimate[2]+1.96 * sqrt(Cop.11$se2),"\n")
cat("T1 dependence: copula param SE:","lb",Cop.2@estimate[1]-1.96 * sqrt(Cop.22$se),
    "estim",Cop.2@estimate[1],"ub",Cop.2@estimate[1]+1.96 * sqrt(Cop.22$se),"\n",
    Cop.2@estimate[2]-1.96 * sqrt(Cop.22$se2),
    "estim",Cop.2@estimate[2],"ub",Cop.2@estimate[2]+1.96 * sqrt(Cop.22$se2),"\n")

cat("T1 dependence: Deviance CN:",-2*t1.result$Cop.1@loglik,"\n")
cat("T1 dependence: Deviance SE:",-2*t1.result$Cop.2@loglik,"\n")
cat("T1 dependence: GoF CN:","test statistics",gof.1$statistic,"p-value",gof.1$p.value,"\n")
cat("T1 dependence: GoF SE:","test statistics",gof.2$statistic,"p-value",gof.2$p.value,"\n")
cat("T1 dependence: Crash prob CN in Q:",Qcrash.1,"\n")
cat("T1 dependence: Crash prob SE in Q:",Qcrash.2,"\n")
cat("T1 dependence: Severe Crash prob CN in Q:",pMvdc(c(x0.1,40),CM0.1),"\n")
cat("T1 dependence: Severe Crash prob SE in Q:",pMvdc(c(x0.1,40),CM0.2),"\n")
cat("T1 dependence: Injury prob CN:",
    Injury.from_cQ_bivariate(t1.result$Cop.1@copula,PIS0,x0.1.un,Qcrash.1*0.3,Conseq.1),"\n")
cat("T1 dependence: Injury prob SE:",
    Injury.from_cQ_bivariate(t1.result$Cop.2@copula,PIS0,x0.2.un,Qcrash.2*0.2,Conseq.2),"\n")

```


## CBPOT nonpar
```{r,echo=FALSE}
source("scripts/CBPOT_kde.R")

scale_factor <- max(plot.dfQ.2$ConditionalD)/max(injury.df$InjuryP)

ggplot(plot.dfQ.2,aes(x=speed,y=ConditionalD)) + 
  geom_line(aes(colour = "SE")) + 
  geom_line(data = plot.dfQ.1,aes(x=speed,y=ConditionalD,colour = "CN")) +
  geom_line(data = injury.df,aes(x=speed,y=InjuryP * scale_factor) ) +
  scale_y_continuous(
    name = "denisty",  # primary y-axis label
    sec.axis = sec_axis(~ . / scale_factor, name = "Injury prob")  # secondary y-axis
  ) +
  scale_colour_manual(name = "Site", values = c("CN" = "red", "SE" = "blue")) +
  labs(title = "Probability density of impact speed",
       x = "speed.u (km/h)", y = "f(y|TTC<0)") +
  theme(panel.grid.major = element_line(colour = "gray91"),
        panel.grid.minor = element_line(colour = "gray88"),
        panel.background = element_rect(fill = "white",
                                        colour = "white", linetype = "solid"),
        plot.background = element_rect(linetype = "solid"))

ggsave("plots/impact_speed_density_TTC_CBPOT_nonpar.png")

v40.un.1 <- pgamma(40,shape = Conseq.1$estimate[1],rate = Conseq.1$estimate[2])
v40.un.2 <- pgamma(40,shape = Conseq.2$estimate[1],rate = Conseq.2$estimate[2])


cat("Nonpar dependence: Severe Crash prob CN in Q:",1- v40.un.1 - x0.1.un + 
      pkdecop(c(x0.1.un,v40.un.1),CopNonpar.1),"\n")
cat("Nonpar dependence: Severe Crash prob SE in Q:",1 - v40.un.2 -x0.2.un +
      pkdecop(c(x0.2.un,v40.un.2),CopNonpar.2),"\n")

Injury.from_cQ_bivariate.nonpar(CopNonpar.1, PIS0,x0.1.un,Qcrash.1*0.3, Conseq.1)
Injury.from_cQ_bivariate.nonpar(CopNonpar.2, PIS0,x0.2.un,Qcrash.2*0.2, Conseq.2)

```


## Injury severity model
Since the focus is on VRU injury, car-related co-variates such as position of the passenger and seat-belt use are of less interest. Based on the Vision Zero principle that the injury occurs when the energy released in the crash is higher human body tolerence, we decide to include only two covariates in the injury severity model -- age and impact speed. Let $Z, Y$ denote the age and impact speed of VRU respectively at a crash, the injury probability given that a crash between VRU and car occurs is given by: 
$$P(S_i|X\leq 0) = \int P(S_i|Z=z,Y=y\})\mathrm{d}F_Z(z|X\leq 0)\mathrm{d}F_Y(y|X\leq 0)$$

This integral can be written in the form $E_Z\big(E_Y(h(Z,Y))\big)$ can be computed numerically. The rest of the theory section is devoted to the identification of a suitable injury severity model for VRU and the derivation of the conditional density $\mathrm{d}F_Y(y|X\leq 0)$ when the crash is extrapolated by extreme value theory.



