---
title: "Draft"
output: html_notebook
---

# loading and creating dataset for modeling
The following scripts loads the data set from SE and CN sites and segment the data set into vehicle-vehicle interaction and vehicle-vru interactions. For the CN dataset, we manually assign the interaction ID so that only interaction where all SMoS measurements are available are kept. For each type of the interaction, datasets for running both BM and POT is created.
```{r,echo=FALSE}
source('preprocessing.R')
```

# Pre-processing Comparision of descriptive statstics
First of all, the interactions with measurements in all TTC, PET and MD are identified and stacked to create the data set for modeling. The histogram and kernel density of the two datasets are ploted below. We will use only TTC and PET in the analysis. We manually removed interaction where delta-V value is greater than 16. From the perspective of computing the injury probability, we think that this is justified as Delta-V at 16 m/s already represents probability of MAIS 3+ injury almost equal to 1. From the data processing view, such interactions often involves a bussess and the tracking of the busses is not accurate because of the size of the tracking box. (2) during the interaction between bus and VRU, the Delta-V extracted is too big (considering the assumed mass of the bus is 50 times of the assumed mass of a human) and creates outlier in fitting the marginal distribution of Delta-V.
```{r,echo=FALSE}
source("descriptive.R")
```

# Univariate model
Block maxima were extracted at 30-minute intervals. To ensure model convergence, we additionally filtered out observations with excessively large values. Data scarcity is a concern at both the CN and SE sites. In some 30-minute blocks, only a few interactions were recorded—none of which were safety-relevant according to SMoS criteria. These sparse and unrepresentative blocks introduce outliers into the sample, leading to convergence issues in the estimation process. This is also why we do not choose r-largest order statistics approach. For successful convergence, we exclude observations which has TTC and PET greater than 4 in the SE dataset. In the CN dataset, we exclude observations which has TTC and PET values greater than 3. The difference in exclusion threshold is motivated by the comparison from Figure 1. 

We start with the vanilla model fitted to the CN and SE sites separately.

```{r,echo=FALSE}

```

# Theory
The injury probability is given by the following mixture model:
$$P(\text{injury}|\text{Crash}) = \int_{z_1,\cdots,z_p} P(\text{injury}|Z_1=\xi_1,...,Z_n=\xi_p)\mathrm{d}F_1(\xi_1)...\mathrm{d}F_p(\xi_p)$$
where $P(\text{injury}|Z_1=\xi_1,...,Z_n=\xi_p)$ accounts for a generalized linear model for injury severity with $Z_i, i=1,...,p$ being the covariates. and $\mathrm{d}F_i(\xi_i)$ is the probability density/mass function of the $i$ th predictor given that a crash occurs. This setup allows injury severity model eitehr specified as logistic regression or multinomial regression. 

To estimate the injury probability at the given site, we need to provide a local (microscopic) definition of crash. This is done by leveraging the SMoS indicator of proximity. Let $X$ denote a proximial SMoS indicator (TTC, PET, etc), then crash is defined as $\{X\leq 0\}$. 

Suppoes $S=\{S_1,...,S_n\}$ is a discrete random variable representing the injury severity, then the probability of the $i$ th injury severity is given by:
$$P(S_i|X\leq 0) = \int_{z_1,\cdots,z_p} P(S_i|\cap_{j=1}^p \{Z_i=\xi_i\})\prod_{j=1}^p\mathrm{d}F_j(\xi_j)$$

depending on the data structure, it may be difficult to quantify the probability density of some predictors $\mathrm{d}F_i(\xi_i)$ due to the absence of data at the given site. For example, if $Z_i$ represent the age of the victim, then $\mathrm{d}F_i(\xi_i)=f(\xi_i|Z\leq 0)\mathrm{d}\xi_i$ is interpreted as the conditional likelihood of age $\xi_i$ given that a crash occurs. Statistics from single site does not provide sufficient information. In this case, one can simulate from different parametric families. 

Since the focus is on VRU injury, car-related co-variates such as position of the passenger and seat-belt use are of less interest. Based on the Vision Zero principle that the injury occurs when the energy released in the crash is higher human body tolerence, we decide to include only two covariates in the injury severity model -- age and impact speed. Let $Z, Y$ denote the age and impact speed of VRU respectively at a crash, the injury probability given that a crash between VRU and car occurs is given by: 
$$P(S_i|X\leq 0) = \int P(S_i|Z=z,Y=y\})\mathrm{d}F_Z(z|X\leq 0)\mathrm{d}F_Y(y|X\leq 0)$$

This integral can be written in the form $E_Z\big(E_Y(h(Z,Y))\big)$ can be computed numerically. The rest of the theory section is devoted to the identification of a suitable injury severity model for VRU and the derivation of the conditional density $\mathrm{d}F_Y(y|X\leq 0)$ when the crash is extrapolated by extreme value theory.

## Injury severity model
```{r,echo=FALSE}}
# AIS 3+ for pedetrain Kong and Yang (2010)
PIS0 <- function(speed){return(1/(1+exp(5.261 - 0.104*speed))) } # Eq.6 
PIS1 <- function(speed,age){return(1/(1+exp(5.15 - 0.101*speed - 0.042*age))) } # Eq.10 

# MAIS 3+ for VRU Lubbe et al (2022)
PIS2 <- function(speed,age){return(1/(1+exp(6.19 - 0.078*speed - 0.038*age))) } # Table 2, for ped
PIS3 <- function(speed,age){return(1/(1+exp(7.47 - 0.079*speed - 0.047*age))) } # Table 2, for cyclist

```

```{r,echo=FALSE}

```

