---
title: "Draft"
output: html_notebook
---

# loading and creating dataset for modeling
The following scripts loads the data set from SE and CN sites and segment the data set into vehicle-vehicle interaction and vehicle-vru interactions. For the CN dataset, we manually assign the interaction ID so that only interaction where all SMoS measurements are available are kept. For each type of the interaction, datasets for running both BM and POT is created.
```{r,echo=FALSE}
source('preprocessing.R')
```

# TTC and Speed_TTC 
```{r,echo=FALSE}
Dat.SE <- SE_TTC
Dat.CN <- CN_TTC_A
x0.1 <- 0
x0.2 <- 0   # crash boundary 

source("BPOT.R")
# estimate
M.1$estimate
confint(M.1)
M.2$estimate
confint(M.2)


# compute probability
pbTvevd(q1=0,q2=40,dep=M.1$estimate[5],thres=c(u1,v1),model="log",
        eta=M.1$nat[1]/M.1$n,
        mar1=c(M.1$estimate[1],M.1$estimate[2]),
        mar2=c(M.1$estimate[3],M.1$estimate[4]),tail.type=2)

# SE TTC model
pbTvevd(q1=0,q2=40,dep=M.2$estimate[5],thres=c(u2,v2),model="log",
        eta=M.2$nat[1]/M.2$n,
        mar1=c(M.2$estimate[1],M.2$estimate[2]),
        mar2=c(M.2$estimate[3],M.2$estimate[4]),tail.type=2)

# plots
# prob distr of impact speed
ggplot(plot.df.1T,aes(x=speed,y=ConditionP)) + 
  geom_line(aes(colour = "CN")) + 
  geom_line(data = plot.df.2T,aes(x=speed,y=ConditionP,colour = "SE")) +
  geom_vline(xintercept = v1, linetype = "dashed", color = "red") +
  geom_vline(xintercept = v2, linetype = "dashed", color = "blue") +
  annotate("text", x = 12, y = 0.5, 
            label = paste("u=", round(v2,3)), color = "blue") +
  annotate("text", x = 20, y = 0.5, 
            label = paste("u=", round(v1,3)), color = "red") +
  scale_colour_manual(name = "Site", values = c("CN" = "red", "SE" = "blue")) +
  labs(title = "Probability distribution of impact speed",
       x = "Speed (km/h)", y = "P(Y <= y | TTC<0)") +
  theme_minimal()

# prob density of impact speed
ggplot(plot.df.1T,aes(x=speed,y=ConditionalD)) + 
  geom_line(aes(colour = "CN")) + 
  geom_line(data = plot.df.2T,aes(x=speed,y=ConditionalD,colour = "SE")) +
  geom_vline(xintercept = v1, linetype = "dashed", color = "red") +
  geom_vline(xintercept = v2, linetype = "dashed", color = "blue") +
  # annotate("text", x = 12, 
  #           label = paste("u=", round(v2,3)), color = "blue") +
  # annotate("text", x = 20, 
  #           label = paste("u=", round(v1,3)), color = "red") +
  scale_colour_manual(name = "Site", values = c("CN" = "red", "SE" = "blue")) +
  labs(title = "Probability density of impact speed",
       x = "Speed (km/h)", y = "f(y|TTC<0)") +
  theme_minimal()

# injury prob
PIS0 <- function(speed){return(1/(1+exp(5.261 - 0.104*speed))) } # Eq.6 
injury.df <- data.frame(speed = seq(0,80,0.5)) %>%
  mutate(InjuryP = sapply(speed, PIS0))

scale_factor <- max(plot.df.1T$ConditionalD)/max(injury.df$InjuryP)

ggplot(plot.df.1T,aes(x=speed,y=ConditionalD)) + 
  geom_line(aes(colour = "CN")) + 
  geom_line(data = plot.df.2T,aes(x=speed,y=ConditionalD,colour = "SE")) +
  geom_line(data = injury.df,aes(x=speed,y=InjuryP * scale_factor) ) +
  geom_vline(xintercept = v1, linetype = "dashed", color = "red") +
  geom_vline(xintercept = v2, linetype = "dashed", color = "blue") +
  scale_y_continuous(
    name = "denisty",  # primary y-axis label
    sec.axis = sec_axis(~ . / scale_factor, name = "Injury prob")  # secondary y-axis
  ) +
  scale_colour_manual(name = "Site", values = c("CN" = "red", "SE" = "blue")) +
  labs(title = "Probability density of impact speed",
       x = "Speed (km/h)", y = "f(y|TTC<0)") +
  theme_minimal()

Injury.from_c_bivariate(plot.df.1T,EVmodel=M.1,Pcrash = Pcrash.1,severity = PIS0)#CN
Injury.from_c_bivariate(plot.df.2T,EVmodel=M.2,Pcrash = Pcrash.2,severity = PIS0)#SE

```

# iTTC and Speed_TTC
```{r,echo=FALSE}
Dat.SE <- SE_iTTC
Dat.CN <- CN_iTTC_A
x0.1 <- 1
x0.2 <- 1
source("BPOT.R")
# estimate
M.1$estimate
confint(M.1)
M.2$estimate
confint(M.2)


# compute probability
pbTvevd(q1=x0.1,q2=40,dep=M.1$estimate[5],thres=c(u1,v1),model="log",
        eta=M.1$nat[1]/M.1$n,
        mar1=c(M.1$estimate[1],M.1$estimate[2]),
        mar2=c(M.1$estimate[3],M.1$estimate[4]),tail.type=2)

# SE TTC model
pbTvevd(q1=x0.2,q2=40,dep=M.2$estimate[5],thres=c(u2,v2),model="log",
        eta=M.2$nat[1]/M.2$n,
        mar1=c(M.2$estimate[1],M.2$estimate[2]),
        mar2=c(M.2$estimate[3],M.2$estimate[4]),tail.type=2)

# plots
# prob distr of impact speed
ggplot(plot.df.1T,aes(x=speed,y=ConditionP)) + 
  geom_line(aes(colour = "CN")) + 
  geom_line(data = plot.df.2T,aes(x=speed,y=ConditionP,colour = "SE")) +
  geom_vline(xintercept = v1, linetype = "dashed", color = "red") +
  geom_vline(xintercept = v2, linetype = "dashed", color = "blue") +
  annotate("text", x = 12, y = 0.5, 
            label = paste("u=", round(v2,3)), color = "blue") +
  annotate("text", x = 20, y = 0.5, 
            label = paste("u=", round(v1,3)), color = "red") +
  scale_colour_manual(name = "Site", values = c("CN" = "red", "SE" = "blue")) +
  labs(title = "Probability distribution of impact speed",
       x = "Speed (km/h)", y = "P(Y <= y | TTC<0)") +
  theme_minimal()

# prob density of impact speed
ggplot(plot.df.1T,aes(x=speed,y=ConditionalD)) + 
  geom_line(aes(colour = "CN")) + 
  geom_line(data = plot.df.2T,aes(x=speed,y=ConditionalD,colour = "SE")) +
  geom_vline(xintercept = v1, linetype = "dashed", color = "red") +
  geom_vline(xintercept = v2, linetype = "dashed", color = "blue") +
  # annotate("text", x = 12, 
  #           label = paste("u=", round(v2,3)), color = "blue") +
  # annotate("text", x = 20, 
  #           label = paste("u=", round(v1,3)), color = "red") +
  scale_colour_manual(name = "Site", values = c("CN" = "red", "SE" = "blue")) +
  labs(title = "Probability density of impact speed",
       x = "Speed (km/h)", y = "f(y|TTC<0)") +
  theme_minimal()

# injury prob
PIS0 <- function(speed){return(1/(1+exp(5.261 - 0.104*speed))) } # Eq.6 
injury.df <- data.frame(speed = seq(0,80,0.5)) %>%
  mutate(InjuryP = sapply(speed, PIS0))

scale_factor <- max(plot.df.1T$ConditionalD)/max(injury.df$InjuryP)

ggplot(plot.df.1T,aes(x=speed,y=ConditionalD)) + 
  geom_line(aes(colour = "CN")) + 
  geom_line(data = plot.df.2T,aes(x=speed,y=ConditionalD,colour = "SE")) +
  geom_line(data = injury.df,aes(x=speed,y=InjuryP * scale_factor) ) +
  geom_vline(xintercept = v1, linetype = "dashed", color = "red") +
  geom_vline(xintercept = v2, linetype = "dashed", color = "blue") +
  scale_y_continuous(
    name = "denisty",  # primary y-axis label
    sec.axis = sec_axis(~ . / scale_factor, name = "Injury prob")  # secondary y-axis
  ) +
  scale_colour_manual(name = "Site", values = c("CN" = "red", "SE" = "blue")) +
  labs(title = "Probability density of impact speed",
       x = "Speed (km/h)", y = "f(y|TTC<0)") +
  theme_minimal()

Injury.from_c_bivariate(plot.df.1T,EVmodel=M.1,Pcrash = Pcrash.1,severity = PIS0,x0=x0.1)#CN
Injury.from_c_bivariate(plot.df.2T,EVmodel=M.2,Pcrash = Pcrash.2,severity = PIS0,x0=x0.2)#SE
```

# Theory
The injury probability is given by the following mixture model:
$$P(\text{injury}|\text{Crash}) = \int_{z_1,\cdots,z_p} P(\text{injury}|Z_1=\xi_1,...,Z_n=\xi_p)\mathrm{d}F_1(\xi_1)...\mathrm{d}F_p(\xi_p)$$
where $P(\text{injury}|Z_1=\xi_1,...,Z_n=\xi_p)$ accounts for a generalized linear model for injury severity with $Z_i, i=1,...,p$ being the covariates. and $\mathrm{d}F_i(\xi_i)$ is the probability density/mass function of the $i$ th predictor given that a crash occurs. This setup allows injury severity model eitehr specified as logistic regression or multinomial regression. 

To estimate the injury probability at the given site, we need to provide a local (microscopic) definition of crash. This is done by leveraging the SMoS indicator of proximity. Let $X$ denote a proximial SMoS indicator (TTC, PET, etc), then crash is defined as $\{X\leq 0\}$. 

Suppoes $S=\{S_1,...,S_n\}$ is a discrete random variable representing the injury severity, then the probability of the $i$ th injury severity is given by:
$$P(S_i|X\leq 0) = \int_{z_1,\cdots,z_p} P(S_i|\cap_{j=1}^p \{Z_i=\xi_i\})\prod_{j=1}^p\mathrm{d}F_j(\xi_j)$$

depending on the data structure, it may be difficult to quantify the probability density of some predictors $\mathrm{d}F_i(\xi_i)$ due to the absence of data at the given site. For example, if $Z_i$ represent the age of the victim, then $\mathrm{d}F_i(\xi_i)=f(\xi_i|Z\leq 0)\mathrm{d}\xi_i$ is interpreted as the conditional likelihood of age $\xi_i$ given that a crash occurs. Statistics from single site does not provide sufficient information. In this case, one can simulate from different parametric families. 



## Injury severity model
Since the focus is on VRU injury, car-related co-variates such as position of the passenger and seat-belt use are of less interest. Based on the Vision Zero principle that the injury occurs when the energy released in the crash is higher human body tolerence, we decide to include only two covariates in the injury severity model -- age and impact speed. Let $Z, Y$ denote the age and impact speed of VRU respectively at a crash, the injury probability given that a crash between VRU and car occurs is given by: 
$$P(S_i|X\leq 0) = \int P(S_i|Z=z,Y=y\})\mathrm{d}F_Z(z|X\leq 0)\mathrm{d}F_Y(y|X\leq 0)$$

This integral can be written in the form $E_Z\big(E_Y(h(Z,Y))\big)$ can be computed numerically. The rest of the theory section is devoted to the identification of a suitable injury severity model for VRU and the derivation of the conditional density $\mathrm{d}F_Y(y|X\leq 0)$ when the crash is extrapolated by extreme value theory.
```{r,echo=FALSE}}
# AIS 3+ for pedetrain Kong and Yang (2010)
PIS0 <- function(speed){return(1/(1+exp(5.261 - 0.104*speed))) } # Eq.6 
PIS1 <- function(speed,age){return(1/(1+exp(5.15 - 0.101*speed - 0.042*age))) } # Eq.10 

# MAIS 3+ for VRU Lubbe et al (2022)
PIS2 <- function(speed,age){return(1/(1+exp(6.19 - 0.078*speed - 0.038*age))) } # Table 2, for ped
PIS3 <- function(speed,age){return(1/(1+exp(7.47 - 0.079*speed - 0.047*age))) } # Table 2, for cyclist

```


