---
title: "R Notebook"
output: html_notebook
---

# Demostration of computing injury probability

## Using bivariate POT methods
First we fit a bivariate POT model using the evd package. The thresholds are chosen according to the `bvtcplot` function
```{r,echo=FALSE}
# 1 stands for CN data. 2 stands for SE data; T stands for TTC; P stands for PET; ST stands for Speed at TTC; SP stands for Speed at PET
thres.order1T <- bvtcplot(CN_TTC_A)$k
thres.order2T <- bvtcplot(SE_TTC)$k
thres.order1P <- bvtcplot(CN_PET_A)$k
thres.order2P <- bvtcplot(SE_PET)$k

u.T1 <- sort(CN_TTC_A$N_TTC,decreasing = TRUE)[thres.order1T]
u.ST1 <- sort(CN_TTC_A$Speed_TTC,decreasing = TRUE)[thres.order1T]
u.T2 <- sort(SE_TTC$N_TTC,decreasing = TRUE)[thres.order2T]
u.ST2 <- sort(SE_TTC$Speed_TTC,decreasing = TRUE)[thres.order2T]

u.P1 <- sort(CN_PET_A$N_PET,decreasing = TRUE)[thres.order1P]
u.SP1 <- sort(CN_PET_A$Speed_PET,decreasing = TRUE)[thres.order1P]
u.P2 <- sort(SE_PET$N_PET,decreasing = TRUE)[thres.order2P]
u.SP2 <- sort(SE_PET$Speed_PET,decreasing = TRUE)[thres.order2P]

M.1T <- fbvpot(x = CN_TTC_A,model = "log",threshold = c(u.T1,u.ST1))
M.1T$estimate
M.2T <- fbvpot(x = SE_TTC,model = "log",threshold = c(u.T2,u.ST2))
M.2T$estimate
M.1P <- fbvpot(x = CN_PET_A,model = "log",threshold = c(u.P1,u.SP1))
M.1P$estimate
M.2P <- fbvpot(x = SE_PET,model = "log",threshold = c(u.P2,u.SP2))
M.2P$estimate

```

We also check if the thresholds gives a good fit in the margin
```{r,echo=FALSE}
prox.1T <- fevd(x = N_TTC,data=CN_TTC_A,threshold = u.T1,period.basis = "two week",
               time.units = "0.5/month", type = "GP")
plot(prox.1T)
prox.1P <- fevd(x = N_PET,data=CN_PET_A,threshold = u.P1,period.basis = "two week",
               time.units = "0.5/month", type = "GP")
plot(prox.1P)
conseq.1T <- fevd(x = Speed_TTC,data=CN_TTC_A,threshold = u.ST1,period.basis = "two week",
                 time.units = "0.5/month", type = "GP")
plot(conseq.1T)
conseq.1P <- fevd(x = Speed_PET,data=CN_PET_A,threshold = u.SP1,period.basis = "two week",
                 time.units = "0.5/month", type = "GP")
plot(conseq.1P)

prox.2T <- fevd(x = N_TTC,data=SE_TTC,threshold = u.T2,period.basis = "month",
               time.units = "months", type = "GP")
plot(prox.2T)
prox.2P <- fevd(x = N_PET,data=SE_PET,threshold = u.P2,period.basis = "month",
               time.units = "months", type = "GP")
plot(prox.2P)
conseq.2T <- fevd(x = Speed_TTC,data=SE_TTC,threshold = u.ST2,period.basis = "month",
                 time.units = "months", type = "GP")
plot(conseq.2T)
conseq.2P <- fevd(x = Speed_PET,data=SE_PET,threshold = u.SP2,period.basis = "month",
                 time.units = "months", type = "GP")
plot(conseq.2P)
```

The probability of a severy crash is computed by truncated bivariate extreme value distribution, such that:
$$P(X>0,Y\leq y) = P(Y\leq y) - P(X\leq 0, Y\leq y)= P(Y\leq y) - G(T(0),T(y))$$

here a logistic model is used, thus $G(z_1,z_2) = \exp(-(z_1^{1/r} + z_2^{1/r})^r)$, where $r$ is the dependence parameter and $z_i = T(x_i)= -\ln(F_i(x_i))$ transforms the margin to standard exponential distribution. $F_i(x_i)=1-P(X_i>u_i)\cdot (1+\gamma_i\cdot \frac{x_i-u_i}{\sigma_i})^{-1/\gamma_i}$ is the unconditional marginal distribution of $X_i$ approximated by EVT.

```{r,echo=FALSE}
# CN TTC model
pbTvevd(q1=0,q2=40,dep=M.1T$estimate[5],thres=c(u.T1,u.ST1),model="log",
        eta=thres.order1T/dim(CN_TTC_A)[1],
        mar1=c(M.1T$estimate[1],M.1T$estimate[2]),
        mar2=c(M.1T$estimate[3],M.1T$estimate[4]),tail.type=4)

# CN PET model
pbTvevd(q1=0,q2=40,dep=M.1P$estimate[5],thres=c(u.P1,u.SP1),model="log",
        eta=thres.order1P/dim(CN_PET_A)[1],
        mar1=c(M.1P$estimate[1],M.1P$estimate[2]),
        mar2=c(M.1P$estimate[3],M.1P$estimate[4]),tail.type=1) # PET margin ends before 0

# SE TTC model
pbTvevd(q1=0,q2=40,dep=M.2T$estimate[5],thres=c(u.T2,u.ST2),model="log",
        eta=thres.order2T/dim(SE_TTC)[1],
        mar1=c(M.2T$estimate[1],M.2T$estimate[2]),
        mar2=c(M.2T$estimate[3],M.2T$estimate[4]),tail.type=4)

# SE PET model
pbTvevd(q1=0,q2=40,dep=M.2P$estimate[5],thres=c(u.P2,u.SP2),model="log",
        eta=thres.order2P/dim(CN_PET_A)[1],
        mar1=c(M.2P$estimate[1],M.2P$estimate[2]),
        mar2=c(M.2P$estimate[3],M.2P$estimate[4]),tail.type=4)  # ends before 0. 
```

We can give the conditional density by computing the following integral:
$$f_Y(y|X\leq 0) = \int_0^{\infty} g(F_X(x),F_Y(y))\mathrm{d}x$$
where $g$ is the joint density of 
```{r,echo=FALSE}
Pcrash.1T <- pevd(0,threshold = u.T1, scale = M.1T$estimate[1],shape = M.1T$estimate[2],
               lower.tail = FALSE,type = "GP") * thres.order1T/dim(CN_TTC_A)[1]
Pcrash.2T <- pevd(0,threshold = u.T2, scale = M.2T$estimate[1],shape = M.2T$estimate[2],
               lower.tail = FALSE,type = "GP") * thres.order2T/dim(SE_TTC)[1]

ss.1T <- seq(u.ST1,60,(60- u.ST1)/150)
ss.2T <- seq(u.ST2,60,(60- u.ST2)/150)

plot.df.1T <- data.frame(speed= ss.1T,
  JointP = sapply(ss.1T,FUN = pbTvevd,q1=0,model="log",dep=M.1T$estimate[5],thres=c(u.T1,u.ST1),
                 eta=thres.order1T/dim(CN_TTC_A)[1],
                 mar1=c(M.1T$estimate[1],M.1T$estimate[2]),
                 mar2=c(M.1T$estimate[3],M.1T$estimate[4]),tail.type=4)) %>%
  mutate(ConditionP = JointP/Pcrash.1T) %>% na.omit()
plot.df.2T <- data.frame(speed= ss.2T,
  JointP = sapply(ss.2T,FUN = pbTvevd,q1=0,model="log",dep=M.2T$estimate[5],thres=c(u.T2,u.ST2),
                 eta=thres.order2T/dim(SE_TTC)[1],
                 mar1=c(M.2T$estimate[1],M.2T$estimate[2]),
                 mar2=c(M.2T$estimate[3],M.2T$estimate[4]),tail.type=4)) %>%
  mutate(ConditionP = JointP/Pcrash.2T)

ggplot(plot.df.1T,aes(x=speed,y=ConditionP)) + 
  geom_line(aes(colour = "CN")) + 
  geom_line(data = plot.df.2T,aes(x=speed,y=ConditionP,colour = "SE")) +
  geom_vline(xintercept = u.ST1, linetype = "dashed", color = "red") +
  geom_vline(xintercept = u.ST2, linetype = "dashed", color = "blue") +
  annotate("text", x = 12, y = 0.5, 
            label = paste("u=", round(u.ST2,3)), color = "blue") +
  annotate("text", x = 20, y = 0.5, 
            label = paste("u=", round(u.ST1,3)), color = "red") +
  scale_colour_manual(name = "Site", values = c("CN" = "red", "SE" = "blue")) +
  labs(title = "Probability distribution of impact speed",
       x = "Speed (km/h)", y = "P(Y <= y | TTC<0)") +
  theme_minimal()

```

Using the conditional probability density, we can compute the injury probability by integrating with the injury severity model. The density of impact speed is plotted below:

```{r,echo=FALSE}
plot.df.1T <- plot.df.1T %>% 
  mutate(ConditionalD = sapply(speed, 
              function(k) c.bivariate(y = k, x = 0, PX = Pcrash.1T, model = "log", 
                      dep = M.1T$estimate[5], thres = M.1T$threshold, 
                      eta = thres.order1T/dim(CN_TTC_A)[1],
                      mar1 = c(M.1T$estimate[1], M.1T$estimate[2]), 
                      mar2 = c(M.1T$estimate[3], M.1T$estimate[4]),ulim=0.45) )
              )
         
plot.df.2T <- plot.df.2T %>% 
  mutate(ConditionalD = sapply(speed, 
              function(k) c.bivariate(y = k, x = 0, PX = Pcrash.2T, model = "log", 
                      dep = M.2T$estimate[5], thres = M.2T$threshold, 
                      eta = thres.order2T/dim(SE_TTC)[1],
                      mar1 = c(M.2T$estimate[1], M.2T$estimate[2]), 
                      mar2 = c(M.2T$estimate[3], M.2T$estimate[4]) ) )
              )

ggplot(plot.df.1T,aes(x=speed,y=ConditionalD)) + 
  geom_line(aes(colour = "CN")) + 
  geom_line(data = plot.df.2T,aes(x=speed,y=ConditionalD,colour = "SE")) +
  geom_vline(xintercept = u.ST1, linetype = "dashed", color = "red") +
  geom_vline(xintercept = u.ST2, linetype = "dashed", color = "blue") +
  # annotate("text", x = 12, 
  #           label = paste("u=", round(u.ST2,3)), color = "blue") +
  # annotate("text", x = 20, 
  #           label = paste("u=", round(u.ST1,3)), color = "red") +
  scale_colour_manual(name = "Site", values = c("CN" = "red", "SE" = "blue")) +
  labs(title = "Probability density of impact speed",
       x = "Speed (km/h)", y = "f(y|TTC<0)") +
  theme_minimal()
```

We use the logistic model for injury severity AIS3+ from kong and Yang (2010):
$$P(AIS3+|Speed) = \frac{1}{1+\exp(5.261 - 0.104*speed)}$$
```{r,echo=FALSE}
# compute injury
# AIS 3+ for pedetrain Kong and Yang (2010)
injury.df <- data.frame(speed = seq(0,60,0.5)) %>%
  mutate(InjuryP = sapply(speed, PIS0))

PIS0 <- function(speed){return(1/(1+exp(5.261 - 0.104*speed))) } # Eq.6 

scale_factor <- max(plot.df.1T$ConditionalD)/max(injury.df$InjuryP)

ggplot(plot.df.1T,aes(x=speed,y=ConditionalD)) + 
  geom_line(aes(colour = "CN")) + 
  geom_line(data = plot.df.2T,aes(x=speed,y=ConditionalD,colour = "SE")) +
  geom_line(data = injury.df,aes(x=speed,y=InjuryP * scale_factor) ) +
  geom_vline(xintercept = u.ST1, linetype = "dashed", color = "red") +
  geom_vline(xintercept = u.ST2, linetype = "dashed", color = "blue") +
  scale_y_continuous(
    name = "denisty",  # primary y-axis label
    sec.axis = sec_axis(~ . / scale_factor, name = "Injury prob")  # secondary y-axis
  ) +
  # annotate("text", x = 12, 
  #           label = paste("u=", round(u.ST2,3)), color = "blue") +
  # annotate("text", x = 20, 
  #           label = paste("u=", round(u.ST1,3)), color = "red") +
  scale_colour_manual(name = "Site", values = c("CN" = "red", "SE" = "blue")) +
  labs(title = "Probability density of impact speed",
       x = "Speed (km/h)", y = "f(y|TTC<0)") +
  theme_minimal()

injury.intergrand <- function(y,EVmodel,Pcrash){
  f.y <- c.bivariate(y = y, x = 0, PX = Pcrash, model = "log", 
                     dep = EVmodel$estimate[5], thres = EVmodel$threshold, 
                     eta = EVmodel$nat[1:2]/dim(SE_TTC)[1],
                     mar1 = c(EVmodel$estimate[1], EVmodel$estimate[2]), 
                     mar2 = c(EVmodel$estimate[3], EVmodel$estimate[4]),ulim=0.45)
  return(PIS0(y) * f.y)
}

#AIS 3+ injury probability | crash
integrate(injury.intergrand,EVmodel=M.1T,Pcrash=Pcrash.1T,
          lower = min(plot.df.1T$speed),upper=max(plot.df.1T$speed))$value
integrate(injury.intergrand,EVmodel=M.2T,Pcrash=Pcrash.2T,lower = min(ss.2T),upper=max(ss.2T))$value


# prob of injury crash
integrate(injury.intergrand,lower = min(ss),upper=Inf)$value * Pcrash

# yearly frequency of injury crash
integrate(injury.intergrand,lower = min(ss),upper=max(ss))$value * Pcrash *
  dim(SE_TTC)[1] * thres.order/dim(SE_TTC)[1] /33 * 365 

# add a plot of conditional density and injury probability (three axis)
```

