---
title: "R Notebook"
output: html_notebook
---

# Demostration of computing injury probability

## Using bivariate POT methods
First we fit a bivariate POT model using the evd package. The thresholds are chosen according to the `bvtcplot` function
```{r,echo=FALSE}
# 1 stands for CN data. 2 stands for SE data; T stands for TTC; P stands for PET; ST stands for Speed at TTC; SP stands for Speed at PET
source("BPOT.R")
M.1T$estimate
M.2T$estimate
M.1P$estimate
M.2P$estimate

```

We also check if the thresholds gives a good fit in the margin
```{r,echo=FALSE}
prox.1T <- fevd(x = N_TTC,data=CN_TTC_A,threshold = u.T1,period.basis = "two week",
               time.units = "0.5/month", type = "GP")
plot(prox.1T)
prox.1P <- fevd(x = N_PET,data=CN_PET_A,threshold = u.P1,period.basis = "two week",
               time.units = "0.5/month", type = "GP")
plot(prox.1P)
conseq.1T <- fevd(x = Speed_TTC,data=CN_TTC_A,threshold = u.ST1,period.basis = "two week",
                 time.units = "0.5/month", type = "GP")
plot(conseq.1T)
conseq.1P <- fevd(x = Speed_PET,data=CN_PET_A,threshold = u.SP1,period.basis = "two week",
                 time.units = "0.5/month", type = "GP")
plot(conseq.1P)

prox.2T <- fevd(x = N_TTC,data=SE_TTC,threshold = u.T2,period.basis = "month",
               time.units = "months", type = "GP")
plot(prox.2T)
prox.2P <- fevd(x = N_PET,data=SE_PET,threshold = u.P2,period.basis = "month",
               time.units = "months", type = "GP")
plot(prox.2P)
conseq.2T <- fevd(x = Speed_TTC,data=SE_TTC,threshold = u.ST2,period.basis = "month",
                 time.units = "months", type = "GP")
plot(conseq.2T)
conseq.2P <- fevd(x = Speed_PET,data=SE_PET,threshold = u.SP2,period.basis = "month",
                 time.units = "months", type = "GP")
plot(conseq.2P)

# add the plot of tail dependence
```

The probability of a severy crash is computed by truncated bivariate extreme value distribution, such that:
$$P(X>0,Y\leq y) = P(Y\leq y) - P(X\leq 0, Y\leq y)= P(Y\leq y) - G(T(0),T(y))$$

here a logistic model is used, thus $G(z_1,z_2) = \exp(-(z_1^{1/r} + z_2^{1/r})^r)$, where $r$ is the dependence parameter and $z_i = T(x_i)= -\ln(F_i(x_i))$ transforms the margin to standard exponential distribution. $F_i(x_i)=1-P(X_i>u_i)\cdot (1+\gamma_i\cdot \frac{x_i-u_i}{\sigma_i})^{-1/\gamma_i}$ is the unconditional marginal distribution of $X_i$ approximated by EVT.

```{r,echo=FALSE}
# CN TTC model
pbTvevd(q1=0,q2=40,dep=M.1T$estimate[5],thres=c(u.T1,u.ST1),model="log",
        eta=thres.order1T/dim(CN_TTC_A)[1],
        mar1=c(M.1T$estimate[1],M.1T$estimate[2]),
        mar2=c(M.1T$estimate[3],M.1T$estimate[4]),tail.type=4)

# CN PET model
pbTvevd(q1=0,q2=40,dep=M.1P$estimate[5],thres=c(u.P1,u.SP1),model="log",
        eta=thres.order1P/dim(CN_PET_A)[1],
        mar1=c(M.1P$estimate[1],M.1P$estimate[2]),
        mar2=c(M.1P$estimate[3],M.1P$estimate[4]),tail.type=1) # PET margin ends before 0

# SE TTC model
pbTvevd(q1=0,q2=40,dep=M.2T$estimate[5],thres=c(u.T2,u.ST2),model="log",
        eta=thres.order2T/dim(SE_TTC)[1],
        mar1=c(M.2T$estimate[1],M.2T$estimate[2]),
        mar2=c(M.2T$estimate[3],M.2T$estimate[4]),tail.type=4)

# SE PET model
pbTvevd(q1=0,q2=40,dep=M.2P$estimate[5],thres=c(u.P2,u.SP2),model="log",
        eta=thres.order2P/dim(CN_PET_A)[1],
        mar1=c(M.2P$estimate[1],M.2P$estimate[2]),
        mar2=c(M.2P$estimate[3],M.2P$estimate[4]),tail.type=4)  # ends before 0. 
```

We can give the conditional density by computing the following integral:
$$f_Y(y|X\leq 0) = \int_{-\infty}^0 g(F_X(x),F_Y(y))\mathrm{d}x$$
where $g$ is the joint density of 
```{r,echo=FALSE}
ggplot(plot.df.1T,aes(x=speed,y=ConditionP)) + 
  geom_line(aes(colour = "CN")) + 
  geom_line(data = plot.df.2T,aes(x=speed,y=ConditionP,colour = "SE")) +
  geom_vline(xintercept = u.ST1, linetype = "dashed", color = "red") +
  geom_vline(xintercept = u.ST2, linetype = "dashed", color = "blue") +
  annotate("text", x = 12, y = 0.5, 
            label = paste("u=", round(u.ST2,3)), color = "blue") +
  annotate("text", x = 20, y = 0.5, 
            label = paste("u=", round(u.ST1,3)), color = "red") +
  scale_colour_manual(name = "Site", values = c("CN" = "red", "SE" = "blue")) +
  labs(title = "Probability distribution of impact speed",
       x = "Speed (km/h)", y = "P(Y <= y | TTC<0)") +
  theme_minimal()

```

Using the conditional probability density, we can compute the injury probability by integrating with the injury severity model. The density of impact speed is plotted below:

```{r,echo=FALSE}

ggplot(plot.df.1T,aes(x=speed,y=ConditionalD)) + 
  geom_line(aes(colour = "CN")) + 
  geom_line(data = plot.df.2T,aes(x=speed,y=ConditionalD,colour = "SE")) +
  geom_vline(xintercept = u.ST1, linetype = "dashed", color = "red") +
  geom_vline(xintercept = u.ST2, linetype = "dashed", color = "blue") +
  # annotate("text", x = 12, 
  #           label = paste("u=", round(u.ST2,3)), color = "blue") +
  # annotate("text", x = 20, 
  #           label = paste("u=", round(u.ST1,3)), color = "red") +
  scale_colour_manual(name = "Site", values = c("CN" = "red", "SE" = "blue")) +
  labs(title = "Probability density of impact speed",
       x = "Speed (km/h)", y = "f(y|TTC<0)") +
  theme_minimal()
```

We use the logistic model for injury severity AIS3+ from kong and Yang (2010):
$$P(AIS3+|Speed) = \frac{1}{1+\exp(5.261 - 0.104*speed)}$$
```{r,echo=FALSE}
# compute injury
# AIS 3+ for pedetrain Kong and Yang (2010)
injury.df <- data.frame(speed = seq(0,80,0.5)) %>%
  mutate(InjuryP = sapply(speed, PIS0))

PIS0 <- function(speed){return(1/(1+exp(5.261 - 0.104*speed))) } # Eq.6 

scale_factor <- max(plot.df.1T$ConditionalD)/max(injury.df$InjuryP)

ggplot(plot.df.1T,aes(x=speed,y=ConditionalD)) + 
  geom_line(aes(colour = "CN")) + 
  geom_line(data = plot.df.2T,aes(x=speed,y=ConditionalD,colour = "SE")) +
  geom_line(data = injury.df,aes(x=speed,y=InjuryP * scale_factor) ) +
  geom_vline(xintercept = u.ST1, linetype = "dashed", color = "red") +
  geom_vline(xintercept = u.ST2, linetype = "dashed", color = "blue") +
  scale_y_continuous(
    name = "denisty",  # primary y-axis label
    sec.axis = sec_axis(~ . / scale_factor, name = "Injury prob")  # secondary y-axis
  ) +
  scale_colour_manual(name = "Site", values = c("CN" = "red", "SE" = "blue")) +
  labs(title = "Probability density of impact speed",
       x = "Speed (km/h)", y = "f(y|TTC<0)") +
  theme_minimal()


```
The injury probability (given a crash) for the two sites are 0.271 (CN) and 0.089 (SE) respectively.
```{r,echo=FALSE,warning=FALSE}
Injury.from_c_bivariate(plot.df.1T,EVmodel=M.1T,Pcrash = Pcrash.1T,severity = PIS0)#CN
Injury.from_c_bivariate(plot.df.2T,EVmodel=M.2T,Pcrash = Pcrash.2T,severity = PIS0)#SE

```


## Using copula
First we identify the conditional marginal distribution $P(X\leq 0|X\leq u)$ using univariate EVT.
```{r,echo=FALSE}
# CN TTC
evmix::mrlplot(CN_TTC_A$N_TTC,try.thresh = quantile(CN_TTC_A$N_TTC, 0.8, na.rm =TRUE))
evmix::tscaleplot(CN_TTC_A$N_TTC, tlim=c(-2.5,-1.5),try.thresh = quantile(CN_TTC_A$N_TTC, 0.8, na.rm =TRUE))
evmix::tshapeplot(CN_TTC_A$N_TTC, tlim=c(-2.5,-1.5),try.thresh = quantile(CN_TTC_A$N_TTC, 0.8, na.rm =TRUE))
v.1T <- quantile(CN_TTC_A$N_TTC, 0.7, na.rm =TRUE)

# CN PET
evmix::mrlplot(CN_PET_A$N_TTC,try.thresh = quantile(CN_PET_A$N_TTC, 0.8, na.rm =TRUE))
evmix::tscaleplot(CN_PET_A$N_TTC, tlim=c(-2.5,-1.5),try.thresh = quantile(CN_PET_A$N_TTC, 0.8, na.rm =TRUE))
evmix::tshapeplot(CN_PET_A$N_TTC, tlim=c(-2.5,-1.5),try.thresh = quantile(CN_PET_A$N_TTC, 0.8, na.rm =TRUE))
v.1P <- quantile(CN_PET_A$N_TTC, 0.8, na.rm =TRUE)

# SE TTC
evmix::mrlplot(SE_TTC$N_TTC,try.thresh = quantile(SE_TTC$N_TTC, 0.8, na.rm =TRUE))
evmix::tscaleplot(SE_TTC$N_TTC, tlim=c(-4,-2.5),try.thresh = quantile(SE_TTC$N_TTC, 0.8, na.rm =TRUE))
evmix::tshapeplot(SE_TTC$N_TTC, tlim=c(-4,-2.5),try.thresh = quantile(SE_TTC$N_TTC, 0.8, na.rm =TRUE))
v.2T <- quantile(SE_TTC$N_TTC, 0.8, na.rm =TRUE)

# SE PET
evmix::mrlplot(SE_PET$N_PET,try.thresh = quantile(SE_PET$N_PET, 0.8, na.rm =TRUE))
evmix::tscaleplot(SE_PET$N_PET, tlim=c(-4,-2),try.thresh = quantile(SE_PET$N_PET, 0.8, na.rm =TRUE))
evmix::tshapeplot(SE_PET$N_PET, tlim=c(-4,-2),try.thresh = quantile(SE_PET$N_PET, 0.8, na.rm =TRUE))
v.2P <- quantile(SE_PET$N_PET, 0.8, na.rm =TRUE)
```

Fitting the conditional marginal distribution to the data.
```{r,echo=FALSE}
POT.1T <- fevd(x = N_TTC,data=CN_TTC_A,threshold = v.1T,period.basis = "month",
               time.units = "0.5/month", type = "GP")
POT.1T$results$par

POT.2T <- fevd(x = N_TTC,data=SE_TTC,threshold = v.2T,period.basis = "month",
               time.units = "months", type = "GP")
POT.2T$results$par

POT.2P <- fevd(x = N_PET,data=SE_PET,threshold = v.2P,period.basis = "month",
               time.units = "months", type = "GP")
POT.2P$results$par # no crash

# finding the parametric distribution for speed | X \leq v
Conseq.1T<- CN_TTC_A %>% subset(N_TTC>v.1T) %>% {.$Speed_TTC}%>% 
  fitdistrplus::fitdist(distr="gamma",method = "mle",start=list(shape=9.5,rate=0.5)) # gamma distribution
Conseq.1T$estimate
plot(Conseq.1T)

Conseq.2T<- SE_TTC %>% subset(N_TTC>v.2T) %>% {.$Speed_TTC}%>% 
  fitdistrplus::fitdist(distr="gamma",method = "mle",start=list(shape=4.8,rate=0.4)) # gamma distribution
Conseq.2T$estimate
plot(Conseq.2T)

Cop.dat.1T <- CN_TTC_A %>% subset(N_TTC>v.1T) %>% 
  dplyr::select(N_TTC,Speed_TTC) %>% 
  mutate(N_TTC = pevd(N_TTC,threshold = v.1T,scale = POT.1T$results$par[1],
                      shape = POT.1T$results$par[2], type = "GP"),
         Speed_TTC = pgamma(Speed_TTC,shape = Conseq.1T$estimate[1],
                      rate = Conseq.1T$estimate[2])) %>% as.matrix()

Cop.dat.2T <- SE_TTC %>% subset(N_TTC>v.2T) %>% 
  dplyr::select(N_TTC,Speed_TTC) %>% 
  mutate(N_TTC = pevd(N_TTC,threshold = v.2T,scale = POT.2T$results$par[1],
                      shape = POT.2T$results$par[2], type = "GP"),
         Speed_TTC = pgamma(Speed_TTC,shape = Conseq.2T$estimate[1],
                      rate = Conseq.2T$estimate[2])) %>% as.matrix()

Cop.1T.gum <- fitCopula(gumbelCopula(dim = 2), data = Cop.dat.1T, method = "mpl")
Cop.1T.gum@estimate

Cop.2T.gum <- fitCopula(gumbelCopula(dim = 2), data = Cop.dat.2T, method = "mpl")
Cop.2T.gum@estimate

Cop.2T.clay <- fitCopula(claytonCopula(dim = 2), data = Cop.dat.2T, method = "mpl")
Cop.2T.clay@estimate

gofCopula(Cop.2T.gum@copula, Cop.dat.2T, method = "Sn", simulation = "pb",
          N = 1000, seed = 1294)
gofCopula(Cop.2T.clay@copula, Cop.dat.2T, method = "Sn", simulation = "pb",
          N = 1000, seed = 1294,verbose=FALSE)

# CM stands for mulitvariate distr with copula dependence structure
CM.1T <- Q.distr.param(Cop.1T.gum@copula,mar1=POT.1T,mar2 = Conseq.1T,type = 4)
pMvdc(c(0,50),CM.1T) # P(X>0,Y<50)

CM.2T <- Q.distr.param(Cop.2T.gum@copula,mar1=POT.2T,mar2 = Conseq.2T,type = 4)
pMvdc(c(0,50),CM.2T) # P(X>0,Y<50)

Qcrash.1T <- pevd(0,threshold = v.1T,scale = POT.1T$results$par[1],
                      shape = POT.1T$results$par[2], type = "GP",lower.tail = FALSE)
Qcrash.2T <- pevd(0,threshold = v.2T,scale = POT.2T$results$par[1],
                      shape = POT.2T$results$par[2], type = "GP",lower.tail = FALSE)

sq.2T <- seq(0,60,0.05)

plot.dfQ.1T <- create_plot.df(sq.2T,x=0,model = CM.1T,PX=Qcrash.1T)
plot.dfQ.2T <- create_plot.df(sq.2T,x=0,model = CM.2T,PX=Qcrash.2T)

Injury.from_cQ_bivariate(plot.dfQ.1T,CM.1T,Qcrash.1T,PIS0)
Injury.from_cQ_bivariate(plot.dfQ.2T,CM.2T,Qcrash.2T,PIS0)
```

Here we plot the conditional density of impact speed given TTC<0
```{r,echo=FALSE}


scale_factor <- max(plot.dfQ.2T$ConditionalD)/max(injury.df$InjuryP)


ggplot(plot.dfQ.2T,aes(x=speed,y=ConditionalD)) + 
  geom_line(aes(colour = "SE")) + 
  geom_line(data = plot.dfQ.1T,aes(x=speed,y=ConditionalD,colour = "CN")) +
  geom_line(data = injury.df,aes(x=speed,y=InjuryP * scale_factor) ) +
  scale_y_continuous(
    name = "denisty",  # primary y-axis label
    sec.axis = sec_axis(~ . / scale_factor, name = "Injury prob")  # secondary y-axis
  ) +
  scale_colour_manual(name = "Site", values = c("CN" = "red", "SE" = "blue")) +
  labs(title = "Probability density of impact speed",
       x = "Speed (km/h)", y = "f(y|TTC<0)") +
  theme_minimal()

```


```{r,echo=FALSE}
# plot the conditionalD at crash and the marginal Q density of Y
ggplot() + 
  geom_line(data=plot.dfQ.2T,aes(x=speed,y=ConditionalD,colour = "CN")) +
  geom_line(data=plot.df.1T,aes(x=speed,y=ConditionalD,colour = "CN")) + 
  geom_line(data = plot.df.2T,aes(x=speed,y=ConditionalD,colour = "SE"))+
  geom_vline(xintercept = u.ST1, linetype = "dashed", color = "red") +
  geom_vline(xintercept = u.ST2, linetype = "dashed", color = "blue") +
  scale_colour_manual(name = "Site", values = c("CN" = "red", "SE" = "blue")) +
  labs(title = "Probability density of impact speed",
       x = "Speed (km/h)", y = "f(y|TTC<0)") +
  theme_minimal()




# plot the conditionalD at crash and the marginal Q given different dep value

CM.2T <- Q.distr.param(Cop.2T.gum@copula,mar1=POT.2T,mar2 = Conseq.2T,type = 4)
CM.2T1 <- Q.distr.param(gumbelCopula(param=2),mar1=POT.2T,mar2 = Conseq.2T,type = 4)
CM.2T2 <- Q.distr.param(gumbelCopula(param = 4),mar1=POT.2T,mar2 = Conseq.2T,type = 4)
CM.2T3 <- Q.distr.param(gumbelCopula(param = 6),mar1=POT.2T,mar2 = Conseq.2T,type = 4)

theortical_density <-data.frame(speed= sq.2T) %>%
   mutate(`r=2` = sapply(speed,cQ.bivariate,x=0,PX=Qcrash.2T,model = CM.2T1)/
           normalize_cQ.bivariate(0,Qcrash.2T,CM.2T1)) %>% 
   mutate(`r=4` = sapply(speed,cQ.bivariate,x=0,PX=Qcrash.2T,model = CM.2T2)/
           normalize_cQ.bivariate(0,Qcrash.2T,CM.2T2)) %>%
   mutate(`r=6` = sapply(speed,cQ.bivariate,x=0,PX=Qcrash.2T,model = CM.2T3)/
           normalize_cQ.bivariate(0,Qcrash.2T,CM.2T3)) %>%
  na.omit() 

ggplot(data=theortical_density) + 
  geom_line(aes(x=speed,y=`r=2`,colour = "r=2")) +
  geom_line(aes(x=speed,y=`r=4`,colour = "r=4")) +
  geom_line(aes(x=speed,y=`r=6`,colour = "r=6")) +
  geom_line(data = plot.dfQ.2T,aes(x=speed,y=ConditionalD,colour = "from data (r=1.01)"))+
  scale_colour_manual(name = "Site", values = c("from data (r=1.01)"="black","r=2" = "red",
                                                "r=4"= "blue","r=6"="green")) +
  labs(title = "Probability density of impact speed",
       x = "Speed (km/h)", y = "f(y|TTC<0)") +
  theme_minimal()

Injury.from_cQ_bivariate(plot.dfQ.2T,CM.2T1,Qcrash.2T,PIS0)
Injury.from_cQ_bivariate(plot.dfQ.2T,CM.2T2,Qcrash.2T,PIS0)
Injury.from_cQ_bivariate(plot.dfQ.2T,CM.2T3,Qcrash.2T,PIS0)

```

