---
title: 'Sensitivity of crash severity '
output:
  html_document:
    df_print: paged
---

```{r,echo=FALSE,include=FALSE}
source('../scripts/fitBPOT.R')
source('../scripts/fitCBPOT.R')
```

# Sensitivity of crash severity density
```{r,echo=FALSE,warning=FALSE}
# one parameter models
chi_list <- c(0.1,0.25,0.5,0.75,0.95)
sen_log_M <- bpot_theoretical_density(bpot_dep_from_chi(chi_list, "log"),
                                      log_M)
sen_hr_M <- bpot_theoretical_density(bpot_dep_from_chi(chi_list, "hr"), hr_M)
sen_neglog_M <- bpot_theoretical_density(bpot_dep_from_chi(chi_list, "neglog"), neglog_M)

plot_theoretical_density(sen_log_M,include_origin = FALSE,
                         title = "Logistic model",
                         filename="density_sensitivity_logistic.png")
plot_theoretical_density(sen_hr_M,include_origin = FALSE,
                         title = "Husler-Reiss model",
                         filename="density_sensitivity_hr.png")
plot_theoretical_density(sen_neglog_M,include_origin = FALSE,
                         title = "Negative logistic model",
                         filename="density_sensitivity_neglog.png")
```

```{r,echo=FALSE,warning=FALSE}
# two parameters 
chi_list <- c(0.1,0.25,0.5)
asy_left <- c(0.45,0.97)
asy_right <- c(0.97,0.45)
params_left <- Map(c,rep(list(asy_left),length(chi_list)),bpot_dep_from_chi(chi_list, "alog",asy=asy_left))
params_right <- Map(c,rep(list(asy_right),length(chi_list)),bpot_dep_from_chi(chi_list, "alog",asy=asy_right))
sen_alog_M_left <- bpot_theoretical_density(params_left,alog_M)
sen_alog_M_right <- bpot_theoretical_density(params_right,alog_M)

plot_theoretical_density(sen_alog_M_left,include_origin = FALSE,
                         title = "theoretical density of crash severity: alog BEVD",
                         filename="density_sensitivity_alogistic_left.png")
plot_theoretical_density(sen_alog_M_right,include_origin = FALSE,
                         title = "theoretical density of crash severity: alog BEVD",
                         filename="density_sensitivity_alogistic_right.png")

```

The conclusions are:
The crash severity densities are less affected for single parameter models when marginal approach is used to estimate the model. The differences under joint approaches can be larger because of the fluctration in estimated crash probabilities.

For the choice of dependence models in BPOT approach, as long as the asymmetry is accounted for, which model to use is less important.


```{r,echo=FALSE,warning=FALSE}
# CBPOT approach
tau_list <- c(-0.5,-0.25,0.05,0.25,0.5)
frank_params <- iTau(frankCopula(dim=2), tau_list)
clayton_params <- iTau(claytonCopula(dim=2), tau_list)
gaussian_params <- iTau(normalCopula(dim=2), tau_list)
sen_frank_M <- cbpot_theoretical_density(frank_params, "frank",frank_M)
sen_clayton_M <- cbpot_theoretical_density(clayton_params, "clayton",clayton_M)
sen_gaussian_M <- cbpot_theoretical_density(gaussian_params, "normal",gaussian_M)

plot_theoretical_density(sen_frank_M,include_origin = FALSE,
                         title = "frank copula",
                         filename="density_sensitivity_frank.png")
plot_theoretical_density(sen_clayton_M,include_origin = FALSE,
                         title = "Clayton copula",
                         filename="density_sensitivity_clayton.png")
plot_theoretical_density(sen_gaussian_M,include_origin = FALSE,
                         title = "Gaussian copula",
                         filename="density_sensitivity_gaussian.png")
```

# Sensitivity of severe crash probabilities

## BPOT severe-crash probabilities
```{r,echo=FALSE,warning=FALSE}
severity_thresholds <- c(20, 30, 40, 50)
chi_list <- c(0.1, 0.25, 0.5)

asy_left <- c(0.45, 0.97)
asy_right <- c(0.97, 0.45)
alog_params <- function(asy) {
  Map(c, rep(list(asy), length(chi_list)),
      bpot_dep_from_chi(chi_list, "alog", asy = asy))
}

bpot_sensitivity <- list(
  logistic = list(
    result = log_M,
    params = bpot_dep_from_chi(chi_list, "log")
  ),
  husler_reiss = list(
    result = hr_M,
    params = bpot_dep_from_chi(chi_list, "hr")
  ),
  negative_logistic = list(
    result = neglog_M,
    params = bpot_dep_from_chi(chi_list, "neglog")
  ),
  alog_left = list(result = alog_M, params = alog_params(asy_left)),
  alog_right = list(result = alog_M, params = alog_params(asy_right))
)

bpot_severe_prob <- do.call(rbind, lapply(names(bpot_sensitivity), function(model_name) {
  spec <- bpot_sensitivity[[model_name]]
  do.call(rbind, lapply(severity_thresholds, function(boundary) {
    probabilities <- bpot_theoretical_severe(
      params = spec$params,
      result = spec$result,
      severity_boundary = boundary
    )
    data.frame(
      model = model_name,
      dependence = chi_list,
      severity_boundary = boundary,
      estimate = probabilities,
      row.names = NULL
    )
  }))
}))
row.names(bpot_severe_prob) <- NULL

plot_severe_probability <- function(data, boundary, y_label, strength_label,
                                    title_prefix, filename) {
  plot_data <- data[data$severity_boundary == boundary, , drop = FALSE]
  plot_data$model <- factor(plot_data$model, levels = unique(data$model))

  p <- ggplot(plot_data, aes(x = model, y = estimate, colour = model)) +
    geom_point(size = 2) +
    facet_wrap(~ dependence, labeller = labeller(dependence = function(x) {
      paste0(strength_label, " = ", x)
    })) +
    scale_y_log10(labels = scales::label_scientific()) +
    labs(title = paste0(title_prefix, " (", boundary, " km/h)"),
         x = "Dependence model", y = y_label, colour = "Dependence model") +
    theme_minimal()

  save_plot(p, filename, width = 10, height = 6)
  p
}

for (boundary in severity_thresholds) {
  print(plot_severe_probability(
    bpot_severe_prob, boundary, "P(severe crash)", "chi", "BPOT",
    paste0("severe_probability_bpot_", boundary, ".png")
  ))
}
```

## CBPOT severe-crash probabilities
```{r,echo=FALSE,warning=FALSE}
# Clayton only supports non-negative Kendall's tau.
tau_list <- c(0.05, 0.25, 0.5, 0.75)

cbpot_sensitivity <- list(
  frank = list(
    result = frank_M,
    cop_name = "frank",
    params = iTau(frankCopula(dim = 2), tau_list)
  ),
  clayton = list(
    result = clayton_M,
    cop_name = "clayton",
    params = iTau(claytonCopula(dim = 2), tau_list)
  ),
  gaussian = list(
    result = gaussian_M,
    cop_name = "normal",
    params = iTau(normalCopula(dim = 2), tau_list)
  )
)

cbpot_severe_prob <- do.call(rbind, lapply(names(cbpot_sensitivity), function(model_name) {
  spec <- cbpot_sensitivity[[model_name]]
  do.call(rbind, lapply(severity_thresholds, function(boundary) {
    probabilities <- cbpot_theoretical_severe(
      params = spec$params,
      cop_name = spec$cop_name,
      result = spec$result,
      severity_boundary = boundary
    )
    data.frame(
      model = model_name,
      dependence = tau_list,
      severity_boundary = boundary,
      estimate = probabilities,
      row.names = NULL
    )
  }))
}))
row.names(cbpot_severe_prob) <- NULL

for (boundary in severity_thresholds) {
  print(plot_severe_probability(
    cbpot_severe_prob, boundary, "P(severe crash | POT)", "tau", "CBPOT",
    paste0("severe_probability_cbpot_", boundary, ".png")
  ))
}
```


# sensitivity of injury probability
```{r,echo=FALSE}
age_mean_list <- seq(20, 60, by = 1)

injury_sen_log_M <- bpot_theoretical_injury(c(0.6,0.7,0.8,0.9,0.99), log_M, age_mean_list,
                                          filename = "injury_sensitivity_log.csv")
injury_sen_hr_M <- bpot_theoretical_injury(c(0.2, 0.5, 0.8,1.2,1.6), hr_M, age_mean_list,
                                          filename = "injury_sensitivity_hr.csv")
injury_sen_neglog_M <- bpot_theoretical_injury(c(0.05, 0.2, 0.45,0.6,0.85), neglog_M, age_mean_list,
                                          filename = "injury_sensitivity_neglog.csv")

injury_sen_frank_M <- cbpot_theoretical_injury(c(1, 1.2, 1.4,1.6), "frank",
                                       frank_M, age_mean_list,conseq,filename = "injury_sensitivity_frank.csv")
injury_sen_clayton_M <- cbpot_theoretical_injury(c(-0.6,-0.3, -0.05,0.1,1,2.5), "clayton",
                                       clayton_M, age_mean_list,conseq,filename = "injury_sensitivity_clayton.csv")
injury_sen_gaussian_M <- cbpot_theoretical_injury(c(-0.7,-0.4,-0.1, 0.1, 0.4,0.7), "normal",
                                       gaussian_M, age_mean_list,conseq,filename = "injury_sensitivity_gaussian.csv")

plot_theoretical_injury(injury_sen_log_M)
plot_theoretical_injury(injury_sen_hr_M)
plot_theoretical_injury(injury_sen_neglog_M)
plot_theoretical_injury(injury_sen_frank_M)
plot_theoretical_injury(injury_sen_clayton_M)
plot_theoretical_injury(injury_sen_gaussian_M)

```
